Multiple imputation of incomplete multilevel data using Heckman selection models

### 2022-11-10 1.Paper/Paper.Rmd
Update paper

Although it is possible to formally rule out MCAR, [@little1988]  
<!-- Stat med uses the AMA citation style. citations at the end of sentence go after the period: "out MCAR. [@little1988]" , and those at the end of a clause go after the period: "out MCAR, [@little1988]". Inline citations like "the clinical practice research datalink (CRPD) [@herrett2015]" are fine though.  See an example here:
https://onlinelibrary.wiley.com/doi/epdf/10.1002/sim.9221 Cool thanks a lot!! -->
it is impossible to assess whether data are MAR or MNAR as the observable data are not enough to test the assumptions of both mechanisms.[@enders2022]  For this reason, researchers often conveniently assume that data are MAR, or present results that are based on a complete case analysis. In practice, however, unless missingness is artificially introduced by study design, e.g. when a test is only taken on patients with certain characteristics, missingness will often (partially) depend on unobserved information.

The missingness mechanism of a variable may be a mixture of MAR and MNAR, depending on the availability of data that may predict the missing data. Therefore, the MAR mechanism may become more influential than the MNAR mechanism as more information is recorded. <!-- MAR does not really become more likely. It is rather that the MAR mechanism becomes stronger than the MNAR mechanism. I have tried to capture this with the alternative text !! perfect thanks :D-->
For instance, healthcare professionals may decide what type of measurement on indication, and depending on their personal experience or behavior decide whether or not to record the reasons for the measurement.

@@

Therefore, the aim of this paper is to develop a multilevel imputation method for incomplete continuous and binary variables that are MNAR. In section \ref{heckman} we provide an introduction to the Heckman model and its estimation, and we extended to a hierarchical setting. Then in section \ref{methods} we define the main steps of the proposed imputation method. In section \ref{simulation} we provide the settings and results of a Monte Carlo simulation study used to evaluate the performance of the proposed imputation method. In section \ref{example} we illustrate the method using the survey information collected in different sub-districts in Uganda to estimate the prevalence of malaria in children. 
Finally we provide a discussion  in section \ref{discussion} about  the results, limitations and propose future extension of our method.

@@

Here $\beta^O_{i}$ and $\beta^S_{i}$ are $p\times1$ and $q\times1$ coefficient parameter vectors and $\epsilon^O_{i}=(\epsilon^O_{i1},\epsilon^O_{i2},...,\epsilon^O_{in_i})^T$ and $\epsilon^S_{i}=(\epsilon^S_{i1},\epsilon^S_{i2},...,\epsilon^S_{in_i})^T$ are the residual terms vectors for the outcome and selection equations, respectively. Generally the same variables can be used on the matrix of predictor variables $X^O_{i}$ and $X^S_{i}$. However, to avoid multicollinearity problems, [@puhani1997] it is recommended to include in $X^S_{i}$ at least one variable that is not included in the outcome model. [@angrist2001] This variable is commonly known as an exclusion restriction variable, and should only be associated with the selection in the sample $r^*_{ij}$ but not with the actual observation $y^*_{ij}$. 
<!-- Can you give some explanation (1-3 sentences?) on the exclusion and restriction assumptions here? -->

@@

<!-- This sentence was too long and has the verbs all the way at the end, making it hard to read. I suggest to rephrase to three shorter sentences: -->
Briefly, in a first stage $\theta_i$ the cluster specific parameters of the Heckman model are estimated. That is, the parameters of the two equations in each study, the outcome model and the selection model, are estimated separately for each of the $N$ clusters. In the second stage, all $\theta_i$ are combined using a random effects meta-analysis model.

<!-- I wouldn't call the distribution imaginary. It is assumed to be real but latent. Some words we could use: conjectured, latent, presupposed, presumed  -->
In a random effects model, $\theta_i$ parameters are assumed to be drawn independently and identically from a latent distribution of parameters with a  population mean $\theta_m$ and a population variance $\psi$. [@higgins2009] Thus, the  $\theta_i=\theta_m + b_i$ can be specified using $b_i \sim N(0,\psi)$ random effects to allow for between-study heterogeneity in observed data relationships, and between-study heterogeneity in missing patterns.

# Method
\label{methods}
We follow a similar approach proposed by Resche-Rigon and White (2018) [@resche-rigon2018] for multilevel imputation of data. Briefly, their method was developed to impute variables from a hierarchical structure (i.e., when there are samples unit grouped within a cluster or group).
This method involves estimating an outcome equation (describing the relationships of the observed data to the missing variable) separately in each cluster, after which the parameter estimates of that equation are pooled using random effects meta-analysis.

@@

Our imputation approach differs crucially from the previous approach as here we estimate two correlated equations (instead of a single outcome equation) in each cluster, thus obtaining $\theta_i=\{\beta_i^O,\beta_i^S,\sigma_i,\rho_i\}$ parameters from both equations which are then pooled into a $\theta_m$ parameter set at the marginal or population level. Our method is basically a univariate imputation method, but since it is implemented in a Gibbs sampling procedure, it can also be used to impute multiple incomplete variables in a data set.
<!-- I don't understand what you mean here: it is univariate and it can be used to impute multiple variables? I would say these two statements contradict each other. -->

@@

<!-- Can you add a reference to the R package here? -->

@@

#### Fit a meta-analysis model
<!-- This paragraph starts of in the the present tense " we pool ", and later it's past tense " we pooled" etc. Stick to one tense. In applied papers, it is normal to use only past tense. In stats papers it is common to use present tense when describing a (new) method in the methods section. -->
In this step, we pool the parameters $\widehat{\theta_i}$ with a random effects meta-analysis model using only the groups with observable information, i.e., those that are not systematically missing and have sufficient information to estimate the heckman model.  In particular, we pooled the $p$ coefficients of the $\beta^O$ outcome equation and estimated a multivariate random effects meta-analysis model with them, similarly we combined all $q$ coefficient parameters of the $\beta^S$ selection equation. 
<!-- The phrase " the log-transformed parameter of $\sigma$" is a description of $\sigma'$, so it is what is known as a commenting or parenthetical clause. That means you need to put a comma before and after this clause: "... on $\sigma'$, the log-transformed parameter of $\sigma$, ...". Although this article is not for the BMJ, I can recommend reading their house-style for writing: https://www.bmj.com/about-bmj/resources-authors/house-style -->
<!-- Read more here if you like: http://content.principia.edu/teaching-excellence/commenting-with-commas-when-and-where-to-use-them/ -->
We also performed a univariate random effects meta_analysis on $\sigma'$, the log-transformed parameter of $\sigma$, and another on $\rho'$, the fisher-transformed parameter of $\rho$.

The meta-analysis model is performed with the \textbf{mixmeta} function of the R package *mixmeta*, which allows the use of maximum likelihood (ML), restricted maximum likelihood (REML), and moments estimation methods. For the simulation and illustrative study, we used the restricted REML estimation method, which is recommended as it has a good balance between insensitivity and efficiency. [@viechtbauer2005]
<!-- Insensitivity is not generally a good thing. I think you should either explain why it is a good thing here: "... it is insensitive to the true value of the mean of the distribution of random effects and it is efficient" (or something similar, feel free to rephrase) , or simplify it to "... a good balance between unbiasedness and efficiency."  -->

@@

#### Draw the cluster parameters $\theta^*_i$
<!-- Throughout the text, consider the order in which you put the symbol and the text: -->
<!-- Put the symbol after the text if the text is a description of the symbol: "shrunk-cluster parameters $\theta^*_i$" . -->
<!-- Put the text after the symbol if the symbol quantifies the text: "$n$ patients" , or " $N$ clusters" . -->
We draw the shrunk-cluster parameters $\theta^*_i$ for each group $i$ from the following posterior distribution conditional on $\theta^*_m$ and $\psi^*$.
\begin{align*}
\theta^*_i\sim N\left(\frac{ \theta^*_m/\psi^*+\widehat{\theta_i}/\widehat{S_{\theta_i}}}{1/\psi^*+1/\widehat{S_{\theta_i}}},\frac{1}{1/\psi^*+1/\widehat{S_{\theta_i}}}\right)
\end{align*}

As can be seen, the mean and variance of the posterior distribution is a combination between the estimated marginal and cluster-specific parameters. Here the weights on the cluster-specific parameters $\widehat{\theta_i}$ and the marginal parameters $\theta^*_m$ are inversely proportional to the within cluster variance $\widehat{S_{\theta_i}}$ and between clusters variance $\psi^*$. For example, when $\widehat{S_{\theta_i}} < \psi^*$ the mean of the conditional distribution gives more weight to the estimated cluster-specific parameter.
Conversely, when $\widehat{S_{\theta_i}}>\psi^*$, more weight is given to the estimated marginal parameters.  In the case of systematic missingness, it is like considering the within-cluster variance to be infinite ($\widehat{S_{\theta_i}}\to\infty$), then letting all parameters rely only on the marginal estimates. 
<!-- What does then mean in this last part? should it be " and lettting all parameters rely ..." ? -->

@@

Having $\theta^*_i$ the shrunk-cluster parameters vector for each group, we back-transform $\sigma^*$ and $\rho^*$ to the original scale. Then $y^{miss}_{ij}$ the missing values can be drawn from $p(y^{miss}_{ij}|\theta^*_i)$ their predictive distribution given $\theta^*_i$ as follows:

@@

The imputed value of the $y_{ij}^{miss}$ missing observation can be drawn from the conditional expectation of $y_{ij}$ on unobserved measurements:
\begin{align*}
\mu&= E[y_{ij}|r_{ij}=0,{\beta^O_{i}}^*,{\beta^{S}_{i}}^*,{\rho_{i}}^*,{\sigma_i}^*]\\
\mu&= x^O_{ij}{\beta^O_{i}}^*+\rho_{i}^*\sigma_i^*\frac{-\phi(x^{S}_{ij}{\beta^{S}_{i}}^*)}{\Phi(-x^{S}_{ij}{\beta^{S}_i}^*)}\\
y_{ij}^{miss}&\sim N(\mu,{\sigma_i^*}^2)
\end{align*}

@@

The missing $y_{ij}^{miss}$ is drawn from a Bernoulli distribution with $p_{ij}^*$ proportion parameter given by $P[y_{ij}=1|r_{ij}=0]$, the conditional probability that $y_{ij}=1$ given that the measure is unobservable ($r_{ij}=0$), in a bivariate probit model: [@greene2018]
\begin{align*}
p^*_{ij} &= P[y_{ij}=1|r_{ij}=0,{\beta^O_{i}}^*,{\beta^{S}_{i}}^*,{\rho_{i}}^*]\\
p^*_{ij} &=\frac{\Phi_2(x^O_{ij}{\beta^{O}_i}^*,-x^S_{ij}{\beta^{S}_i}^*,-\rho_i)}{\Phi(-x^{S}_{ij}{\beta^{S}_i}^*)}\\
y_{ij}^{miss}&\sim Ber(p^*_{ij})
\end{align*}

@@

#### Aim
We designed a simulation study aimed to compare the performance of imputation methods for imputing a missing variable in a hierarchical dataset, where the missingness follows a MNAR mechanism.

#### Data-generation mechanism

We generated the data from the Heckman selection model with bivariate normal distribution error terms. For simplicity we assume that the database collects information from $N=10$ clusters of equal number of individuals $n_i=1000$. 
<!-- But you also have scenarios with different numbers for N and n_i. You could write here that for the "basic scenario" we have N =  and n_i = , and that we alter this in other scenarios, or in " senstivity analyses" . -->
For each dataset, we generated $X_{1i}$ a treatment indicator variable from a Bernoulli distribution with a probability of treatment on each cluster equal to 0.6. Next, we simulated the mean of two continuous covariates from a multivariate normal distribution $\mu_h\sim N(\begin{psmallmatrix}0\\0\end{psmallmatrix},\begin{psmallmatrix}0.2 & 0.015\\0.015 & 0.2\end{psmallmatrix})$, with $h=\{2,3\}$. We then simulated for each cluster a baseline covariate $X_{2i}\sim N(\mu_2,1)$ and a exclusion restriction  $X_{3i} \sim N(\mu_3,0.5)$.

Here, we considered $X_{1i}$ and $X_{2i}$ as predictors in the outcome equation $X_i^O=[1,X_{1i},X_{2i}]$. For the selection equation we included both variables and the $X_{3i}$ exclusion restriction, $X_i^S=[1,X_{1i},X_{2i},X_{3i}]$. Then in case of a missing continuous variable, we calculate the latent variables  $y_{i}^*$ and $r^*_{i}$ as follows:
\begin{align*}
 y_{i}^*&= \beta_i^OX_i^O+\epsilon^O_{i} \\
 r_{i}^*&= \beta_i^SX_i^O+\epsilon^S_{i}
\end{align*}
Here we assumed that all coefficient parameters varied across studies, by including cluster-specific random effects as:
\begin{align*}
 \beta_{hi}^O&= \beta_h^O + b_{hi}^O\\
 \beta_{hi}^S&= \beta_h^S + b_{hi}^S
\end{align*}
We fixed coefficients $\beta_h^O=\{0.3,1,1\}$  and $\beta_h^S=\{-0.8,1.3,-0.7,1.2\}$ in order to get around 40\% of sporadically missing values on the response $y_{ij}$ in the entire data set. Additionally, we ensured that the $y_{ij}^*$ observations were systematically missing in 20% of the clusters included in the data set. 
<!-- How did we ensure that these were systematically missing in 20% of the clusters? Does that follow from the fixing of the betas? Or did you remove the y values in 20% of the clusters? Please be specific. -->
We assumed that random effects were independent within equations ($b_{h0}^O\indep b_{h1}^O \indep b_{h2}^O$ and $b_{h0}^S \indep b_{h1}^S \indep b_{h2}^S$), but were linked between both selection and outcome equations through a bivariate normal distributed as: 
\[\begin{pmatrix}
b^O_{h} \\
b^S_{h}
\end{pmatrix}\sim N\left(\begin{pmatrix}
0 \\
0
\end{pmatrix},\sigma_{bh}^2\begin{pmatrix}
1 & \rho*0.4 \\
\rho*0.4 &1 
\end{pmatrix}\right)
\]
with the parameters $\sigma_{b0}^2=\sigma_{b1}^2=\sigma_{b2}^2=0.4$.  We considered that the correlation parameter of the random effects between equations is 40% of the value of the assumed correlation parameter between error terms $\rho$. In addition, we included a random effect on the exclusion restriction variable given by $b_3\sim N(0,0.2)$ assuming that the intracluster variation in the exclusion restriction effect is lower than the variation on other coefficient parameters effects. The $\rho$ parameter was given different values depending on the simulated missing mechanism (See below additional scenarios) .

@@

##### Adittional scenarios

@@

- \textbf{M(N)AR scenarios:} We assessed whether the model performed well in terms of bias and coverage when the data followed a missing MAR mechanism ($\rho=0$), and when it followed a MNAR mechanism with a low ($\rho=0.3$), intermediate ($\rho=0.6$) and strong correlation ($\rho=0.9$) between $y^*$ and $r^*$.
- \textbf{Influence of sample size}: Model sensitivity was analysed with respect to $n_i=\{50,100,1000\}$ the number of patients per cluster and $N=\{10,50,100\}$ the number of studies. We consider the 
<!-- The sentence above stops mid-sentence. -->
- \textbf{Violation of distributional assumptions:} To assess how the imputation models behave in the face of deviations from normality assumptions, we simulated data in which the errors followed a skewed t-distribution and also in which the missing process follows a MNAR mechanism with an explicit truncated model, i.e., the participation is directly related to the value of the outcome variable. 
- \textbf{Binary response}: We evaluated the imputation method when the missing variables are binary. Therefore we simulated  $y_{i}$ binary incomplete variables, we keep the parameters similar to the ones used in the simulation of missing continuous variables, but the observable binary variables were defined as:
\begin{align*}
    r_{i}&=I(r*_{i}>0)\\
    y_{i}&=I(y^*_{i}>0) \forall r^*_{ij}>0
\end{align*}
<!-- One binary variable or multiple binary variables? -->

#### Estimand
The estimands were the parameter coefficients of the outcome equation $\beta^O={\beta^O_0,\beta^O_1,\beta^O_2}$, with special emphasis on the treatment effect parameter $\beta^0_1$. We also report the estimated variance of the random effects and residual errors $\sigma_{b0}^2$,$\sigma_{b1}^2$,$\sigma_{b2}^2$,$\sigma_{e}^2$.

After the imputation procedure, we estimated the following (generalized) mixed linear effect model using the *lmer()* function from the *lme4* R package.[@bates2022]
<!-- Add ref to lme4 package -->
$y_{i}=\beta_i^OX_i^O+\epsilon^O_{i}$
In case of missing binary variable, we used the same matrix of predictors but on a binary model estimated with the *glmer()* function from the *lme4* R package. Then, we pooled the estimates of the $\beta_i^O$ and the variance of the random effect and residual errors of the multiple imputed datasets according to Rubin's rule, [@rubin1987] over which we calculated the performance measures on the estimands.

@@

#### Method

@@

#### Performance measures

@@

- \textbf{Bias:} Bias on the coefficient and random effect parameters.
- \textbf{Coverage:} Coverage of the 95\% confidence intervals for the coefficient  parameters.
- \textbf{Width}: Width of the confidence interval on the coefficient parameters.
- \textbf{RMSE:} Root mean squared error of the coefficient  and random effect parameters.

@@

#### Software
For the simulation study and illustrative examples we used R version 4.0.4 in a linux environment. 
<!-- Add a reference to R if you haven't done so yet. Run citation() in R to get a bibtex entry. -->
The Heckman 2L imputation method is available in the mice R package (as \textbf{mice.2l.heckman()}) and also on the github repository \url{https://github.com/johamunoz/Heckman2l} where you can also find all the codes accompanying this paper and a toy example that explains how to implement the method in mice.

#### Results

@@

#### Results

>>>>

<!-- Stat med uses the AMA citation style. citations at the end of sentence go after the period: "out MCAR. [@little1988]" , and those at the end of a clause go after the period: "out MCAR, [@little1988]". Inline citations like "the clinical practice research datalink (CRPD) [@herrett2015]" are fine though.  See an example here:
https://onlinelibrary.wiley.com/doi/epdf/10.1002/sim.9221 Cool thanks a lot!! -->
Although it is possible to formally rule out MCAR, [@little1988]  it is impossible to assess whether data are MAR or MNAR as the observable data are not enough to test the assumptions of both mechanisms.[@enders2022]  For this reason, researchers often conveniently assume that data are MAR, or present results that are based on a complete case analysis. In practice, however, unless missingness is artificially introduced by study design, e.g. when a test is only taken on patients with certain characteristics, missingness will often (partially) depend on unobserved information.

The missingness mechanism of a variable may be a mixture of MAR and MNAR, depending on the availability of data that may predict the missing data. Therefore, the MAR mechanism may become more influential than the MNAR mechanism as more information is recorded. <!-- MAR does not really become more likely. It is rather that the MAR mechanism becomes stronger than the MNAR mechanism. I have tried to capture this with the alternative text !! perfect thanks, your point is correct and now sound so much better :D-->
For instance, healthcare professionals may decide what type of measurement on indication, and depending on their personal experience or behavior decide whether or not to record the reasons for the measurement.

@@

Therefore, the aim of this paper is to develop a multilevel imputation method for incomplete continuous and binary variables that are MNAR. In section \ref{heckman} we provide an introduction to the Heckman model and its estimation, and we extended to a hierarchical setting. Then in section \ref{methodsn} we define the main steps of the proposed imputation method. In section \ref{simulation} we provide the settings and results of a Monte Carlo simulation study used to evaluate the performance of the proposed imputation method. In section \ref{example} we illustrate the method using the survey information collected in different sub-districts in Uganda to estimate the prevalence of malaria in children. 
Finally we provide a discussion  in section \ref{discussion} about  the results, limitations and propose future extension of our method.

@@

Here $\beta^O_{i}$ and $\beta^S_{i}$ are $p\times1$ and $q\times1$ coefficient parameter vectors and $\epsilon^O_{i}=(\epsilon^O_{i1},\epsilon^O_{i2},...,\epsilon^O_{in_i})^T$ and $\epsilon^S_{i}=(\epsilon^S_{i1},\epsilon^S_{i2},...,\epsilon^S_{in_i})^T$ are the residual terms vectors for the outcome and selection equations, respectively.

Generally the same variables can be used on the matrix of predictor variables $X^O_{i}$ and $X^S_{i}$. However, to avoid multicollinearity problems, [@puhani1997] it is recommended to include in $X^S_{i}$ at least one variable that is not included in the outcome model. [@angrist2001] This variable is commonly known as an exclusion restriction variable (ERV), and should only be associated with the selection in the sample $r^*_{ij}$ (relevance condition) but not with the actual observation $y^*_{ij}$ (exclusion condition).[@gomes_etal20] The ERV should meet the relevance and exclusion conditions in order to provide independent information about the selection processing in order to facilitate the estimation of the Heckman model specially when dataset does not follow the BVN assumption.
<!-- Can you give some explanation (1-3 sentences?) on the exclusion and restriction assumptions here? -->

@@

<!-- This sentence was too long and has the verbs all the way at the end, making it hard to read. I suggest to rephrase to three shorter sentences: OK!!-->
Briefly, in a first stage  we estimate $\theta_i$ the cluster specific parameters of the Heckman model. That is, we estimate separately for each of the $N$ clusters,the parameters of the two equations in each study, the outcome model and the selection model. In the second stage, we combine all $\theta_i$  by using a random effects meta-analysis model.

<!-- I wouldn't call the distribution imaginary. It is assumed to be real but latent. Some words we could use: conjectured, latent, presupposed, presumed, Thanks :D -->
In a random effects model, $\theta_i$ parameters are assumed to be drawn independently and identically from a latent distribution of parameters with a  population mean $\theta_m$ and a population variance $\psi$. [@higgins2009] Thus, the  $\theta_i=\theta_m + b_i$ can be specified using $b_i \sim N(0,\psi)$ random effects to allow for between-study heterogeneity in observed data relationships, and between-study heterogeneity in missing patterns.

# Method
\label{methodsn}
We follow a similar approach proposed by Resche-Rigon and White (2018) [@resche-rigon2018] for multilevel imputation of data. Briefly, their method was developed to impute variables from a hierarchical structure (i.e., when there are samples unit grouped within a cluster or group).
This method involves estimating an outcome equation (describing the relationships of the observed data to the missing variable) separately in each cluster, after which the parameter estimates of that equation are pooled using random effects meta-analysis.

@@

Our imputation approach differs crucially from the previous approach as here we estimate two correlated equations (instead of a single outcome equation) in each cluster, thus obtaining $\theta_i=\{\beta_i^O,\beta_i^S,\sigma_i,\rho_i\}$ parameters from both equations which are then pooled into a $\theta_m$ parameter set at the marginal or population level. Our imputation method was created for impute univariate incomplete datasets, but since it is implemented in a Gibbs sampling procedure, it can also be used to impute multiple incomplete variables in a data set.
<!-- I don't understand what you mean here: it is univariate and it can be used to impute multiple variables? I would say these two statements contradict each other. yes you are right i tried to say is a mehtod for univariate incomplete dataset -->

@@

<!-- Can you add a reference to the R package here? OK -->

@@

#### Fit a meta-analysis model
<!-- This paragraph starts of in the the present tense " we pool ", and later it's past tense " we pooled" etc. Stick to one tense. In applied papers, it is normal to use only past tense. In stats papers it is common to use present tense when describing a (new) method in the methods section. -->
In this step, we pool the parameters $\widehat{\theta_i}$ with a random effects meta-analysis model using only the groups with observable information, i.e., those that are not systematically missing and have sufficient information to estimate the heckman model.  In particular, we pool the $p$ coefficients of the $\beta^O$ outcome equation and estimated a multivariate random effects meta-analysis model with them, similarly we combine all $q$ coefficient parameters of the $\beta^S$ selection equation. 
<!-- The phrase " the log-transformed parameter of $\sigma$" is a description of $\sigma'$, so it is what is known as a commenting or parenthetical clause. That means you need to put a comma before and after this clause: "... on $\sigma'$, the log-transformed parameter of $\sigma$, ...". Although this article is not for the BMJ, I can recommend reading their house-style for writing: https://www.bmj.com/about-bmj/resources-authors/house-style -->
<!-- Read more here if you like: http://content.principia.edu/teaching-excellence/commenting-with-commas-when-and-where-to-use-them/ Thanks a lot you are a walking library :D, very appreciated all your lessons!! -->
We also performed a univariate random effects meta_analysis on $\sigma'$, the log-transformed parameter of $\sigma$, and another on $\rho'$, the fisher-transformed parameter of $\rho$.

The meta-analysis model is performed with the \textbf{mixmeta} function of the R package *mixmeta*, which allows the use of maximum likelihood (ML), restricted maximum likelihood (REML), and moments estimation methods. For the simulation and illustrative study, we used the restricted REML estimation method, which is recommended as it has a good balance between unbiasedness and efficiency. [@viechtbauer2005]
<!-- Insensitivity is not generally a good thing. I think you should either explain why it is a good thing here: "... it is insensitive to the true value of the mean of the distribution of random effects and it is efficient" (or something similar, feel free to rephrase) , or simplify it to "... a good balance between unbiasedness and efficiency. jajaj I think at certain point i used the translator to correct my writing and i did not notice that change unbiasedness by insensitive..But thanks God you have eagle eyes @@) .."  -->

@@

#### Draw the cluster parameters $\theta^*_i$
<!-- Throughout the text, consider the order in which you put the symbol and the text: -->
<!-- Put the symbol after the text if the text is a description of the symbol: "shrunk-cluster parameters $\theta^*_i$" . -->
<!-- Put the text after the symbol if the symbol quantifies the text: "$n$ patients" , or " $N$ clusters" . Whoahh you are a gramatical Lord thanks a lot!! -->
We draw the shrunk-cluster parameters $\theta^*_i$ for each group $i$ from the following posterior distribution conditional on $\theta^*_m$ and $\psi^*$.
\begin{align*}
\theta^*_i\sim N\left(\frac{ \theta^*_m/\psi^*+\widehat{\theta_i}/\widehat{S_{\theta_i}}}{1/\psi^*+1/\widehat{S_{\theta_i}}},\frac{1}{1/\psi^*+1/\widehat{S_{\theta_i}}}\right)
\end{align*}

As can be seen, the mean and variance of the posterior distribution is a combination between the estimated marginal and cluster-specific parameters. Here the weights on the cluster-specific parameters $\widehat{\theta_i}$ and the marginal parameters $\theta^*_m$ are inversely proportional to the within cluster variance $\widehat{S_{\theta_i}}$ and between clusters variance $\psi^*$. For example, when $\widehat{S_{\theta_i}} < \psi^*$ the mean of the conditional distribution gives more weight to the estimated cluster-specific parameter.
Conversely, when $\widehat{S_{\theta_i}}>\psi^*$, more weight is given to the estimated marginal parameters. Therefore, in case of a cluster with systematic missingness, it is as if the within-cluster variance is infinite ($\widehat{S_{\theta_i}}\to\infty$), so all the weight is assigned to the parameters estimated at the marginal level. That is, when there is no information in a cluster, we rely entirely on the marginal information.

<!-- What does then mean in this last part? should it be " and lettting all parameters rely ..." ? I think now is more clear, thanks!! -->

@@

Having $\theta^*_i$, the shrunk-cluster parameters vector for each group, we back-transform $\sigma^*$ and $\rho^*$ to the original scale. Then $y^{miss}_{ij}$, the missing values, can be drawn from $p(y^{miss}_{ij}|\theta^*_i)$, their predictive distribution given $\theta^*_i$, as follows:

@@

The imputed value of $y_{ij}^{miss}$ can be drawn from the conditional expectation of $y_{ij}$ on unobserved measurements:
\begin{align*}
\mu&= E[y_{ij}|r_{ij}=0,{\beta^O_{i}}^*,{\beta^{S}_{i}}^*,{\rho_{i}}^*,{\sigma_i}^*]\\
\mu&= x^O_{ij}{\beta^O_{i}}^*+\rho_{i}^*\sigma_i^*\frac{-\phi(x^{S}_{ij}{\beta^{S}_{i}}^*)}{\Phi(-x^{S}_{ij}{\beta^{S}_i}^*)}\\
y_{ij}^{miss}&\sim N(\mu,{\sigma_i^*}^2)
\end{align*}

@@

When $y_{ij}^{miss}$ is a binary variable, the imputed value is drawn from a Bernoulli distribution with a proportion parameter $p_{ij}^*$  given by $P[y_{ij}=1|r_{ij}=0]$, that is the conditional probability that $y_{ij}=1$ given that the measure is unobservable ($r_{ij}=0$). The $p_{ij}^*$ is obtained from a bivariate probit model, as follows: [@greene2018]
\begin{align*}
p^*_{ij} &= P[y_{ij}=1|r_{ij}=0,{\beta^O_{i}}^*,{\beta^{S}_{i}}^*,{\rho_{i}}^*]\\
p^*_{ij} &=\frac{\Phi_2(x^O_{ij}{\beta^{O}_i}^*,-x^S_{ij}{\beta^{S}_i}^*,-\rho_i)}{\Phi(-x^{S}_{ij}{\beta^{S}_i}^*)}\\
y_{ij}^{miss}&\sim Ber(p^*_{ij})
\end{align*}

@@

## Aim
We designed a simulation study aimed to compare the performance of imputation methods for imputing a missing variable in a hierarchical dataset, where the missingness follows a MNAR mechanism.

## Data-generation mechanism

We generated the data from the Heckman selection model with bivariate normal distribution error terms. For simplicity we assume a "basic scenario" where the database collects information from $N=10$ clusters of equal number of individuals $n_i=1000$. However, we altered both $N$ and $n_i$ in the sensitivity analysis (see below). 
<!-- But you also have scenarios with different numbers for N and n_i. You could write here that for the "basic scenario" we have N =  and n_i = , and that we alter this in other scenarios, or in " senstivity analyses" ok . -->
For each dataset, we generated $X_{1i}$ a treatment indicator variable from a Bernoulli distribution with a probability of treatment on each cluster equal to 0.6. Next, we simulated the mean of two continuous covariates from a multivariate normal distribution $\mu_h\sim N(\begin{psmallmatrix}0\\0\end{psmallmatrix},\begin{psmallmatrix}0.2 & 0.015\\0.015 & 0.2\end{psmallmatrix})$, with $h=\{2,3\}$. We then simulated for each cluster a baseline covariate $X_{2i}\sim N(\mu_2,1)$ and a exclusion restriction  $X_{3i} \sim N(\mu_3,0.5)$.

Here, we considered $X_{1i}$ and $X_{2i}$ as predictors in the outcome equation $X_i^O=[1,X_{1i},X_{2i}]$. For the selection equation we included both variables and the $X_{3i}$ exclusion restriction, $X_i^S=[1,X_{1i},X_{2i},X_{3i}]$. Then in case of a missing continuous variable, we calculate the latent variables  $y_{i}^*$ and $r^*_{i}$ as follows:
\begin{align*}
 y_{i}^*&= \beta_i^OX_i^O+\epsilon^O_{i} \\
 r_{i}^*&= \beta_i^SX_i^O+\epsilon^S_{i}
\end{align*}
Here we assumed that all coefficient parameters varied across studies, by including cluster-specific random effects as:
\begin{align*}
 \beta_{hi}^O&= \beta_h^O + b_{hi}^O\\
 \beta_{hi}^S&= \beta_h^S + b_{hi}^S
\end{align*}
We fixed coefficients $\beta_h^O=\{0.3,1,1\}$  and $\beta_h^S=\{-0.8,1.3,-0.7,1.2\}$ in order to get around 40\% of sporadically missing values on the response $y_{ij}$ in the entire data set. Additionally, we ensured that the $y_{ij}^*$ observations were systematically missing in 20% of the clusters included in the data set, by removing the outcome values in the 20% of the clusters. 
<!-- How did we ensure that these were systematically missing in 20% of the clusters? Does that follow from the fixing of the betas? Or did you remove the y values in 20% of the clusters? Please be specific. OK-->
We assumed that random effects were independent within equations ($b_{h0}^O\indep b_{h1}^O \indep b_{h2}^O$ and $b_{h0}^S \indep b_{h1}^S \indep b_{h2}^S$), but were linked between both selection and outcome equations through a bivariate normal distributed as: 
\[\begin{pmatrix}
b^O_{h} \\
b^S_{h}
\end{pmatrix}\sim N\left(\begin{pmatrix}
0 \\
0
\end{pmatrix},\sigma_{bh}^2\begin{pmatrix}
1 & \rho*0.4 \\
\rho*0.4 &1 
\end{pmatrix}\right)
\]
with the parameters $\sigma_{b0}^2=\sigma_{b1}^2=\sigma_{b2}^2=0.4$.  We considered that the correlation parameter of the random effects between equations is 40% of the value of the assumed correlation parameter between error terms $\rho$. In addition, we included a random effect on the exclusion restriction variable given by $b_3\sim N(0,0.2)$ assuming that the intracluster variation in the exclusion restriction effect is lower than the variation on other coefficient parameters effects. The $\rho$ parameter was given different values depending on the simulated missing mechanism (See below additional scenarios) .

@@

## Adittional scenarios

@@

- \textbf{M(N)AR scenarios:} We assessed whether the model performed well in terms of bias and coverage when the data followed a missing MAR mechanism ($\rho=0$), and when data followed a MNAR mechanism with a low ($\rho=0.3$), intermediate ($\rho=0.6$) and strong correlation ($\rho=0.9$) between $y^*$ and $r^*$. We evaluated also the imputation method when the missing variable is binary. Therefore we simulated a $y_{i}$ binary incomplete variable, by keeping similar parameters to the ones used in the simulation of a missing continuous variable, but here we defined the observable binary variable as:
\begin{align*}
    r_{i}&=I(r*_{i}>0)\\
    y_{i}&=I(y^*_{i}>0) \forall r^*_{ij}>0
\end{align*}

- \textbf{Influence of sample size}: Model sensitivity was analysed with respect to the the number of patients per cluster $n_i=\{50,100,1000\}$  and the number of studies $N=\{10,50,100\}$.

- \textbf{Violation of distributional assumptions:} To assess how the imputation models behave in the face of deviations from normality assumptions, we simulated data in which the errors followed a skewed t-distribution and also in which the missing process follows a MNAR mechanism with an explicit truncated model, i.e., the participation is directly related to the value of the outcome variable.

## Estimand
The estimands were the parameter coefficients of the outcome equation $\beta^O={\beta^O_0,\beta^O_1,\beta^O_2}$, with special emphasis on the treatment effect parameter $\beta^0_1$. We also report the estimated variance of the random effects and residual errors $\sigma_{b0}^2$,$\sigma_{b1}^2$,$\sigma_{b2}^2$,$\sigma_{e}^2$.

After the imputation procedure, we estimated the following (generalized) mixed linear effect model using the *lmer()* function from the *lme4* R package.[@bates2022]
$y_{i}=\beta_i^OX_i^O+\epsilon^O_{i}$
In case of missing binary variable, we used the same matrix of predictors but on a binary model estimated with the *glmer()* function from the *lme4* R package. Then, we pooled the estimates of the $\beta_i^O$ and the variance of the random effect and residual errors of the multiple imputed datasets according to Rubin's rule, [@rubin1987] over which we calculated the performance measures on the estimands.

@@

## Method

@@

## Performance measures

@@

- \textbf{Bias:} Bias on the coefficient and random effect parameters.
- \textbf{Coverage:} Coverage of the 95\% confidence intervals for the coefficient  parameters.
- \textbf{Width}: Average width of the confidence interval on the coefficient parameters.
- \textbf{RMSE:} Root mean squared error of the coefficient  and random effect parameters.

@@

## Software
For the simulation study and illustrative examples we used R version 4.0.4 in a linux environment. [@Rpackage]
<!-- Add a reference to R if you haven't done so yet. Run citation() in R to get a bibtex entry. Mago!!! -->
The Heckman 2L imputation method is available in the mice R package (as \textbf{mice.2l.heckman()}) and also on the github repository \url{https://github.com/johamunoz/Statsmed_Heckman} where you can also find all the codes accompanying this paper and a toy example that explains how to implement the method in mice.

## Results

@@

When we used the Heckman model in an explicit MNAR process, we observed that our model might not be completely suitable for this type of scenario, as it highly affected the bias of the intercept parameter ($\beta_0$) and the width of the 95\% CI of all the coefficient parameters \ref{fig:dist}. Also the bias of the random effects parameters ($\sigma_0$ and $\sigma_2$) was affected.

With respect to the Skewed-t scenario, only the bias of $\beta_0$ was affected. For both scenarios, it is also seen that the coverage in the $\beta_0$ was highly affected.
